---
title: "In-class Exercise 3"
author: "Immanuel Leong"
date: "September 2, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# In-class Exercise 3: Spatial Point Patterns Analysis: spatstat methods

```{r}
pacman::p_load(sf, spatstat, raster, tmap, tidyverse)
```

```{r}
childcare_sf <- st_read("data/child-care-services-geojson.geojson")

mpsz_sf <- st_read(dsn = "data", layer = "MP14_SUBZONE_WEB_PL")
```

## Issue 1: Installing mapstools

**maptools** is retired and binary is removed from CRAN. However, we can download it from Posit Public Package Manager using the code chunk below.

```{r}
#| eval: false
# Avoid maptools being repetitively installed every time the Quarto document is rendered

#install.packages("maptools", repos = "https://packagemanager.posit.co/cran/2023-10-13")
```

## Issue 2: Creating coastal outline

*st_combine()* returns a single, combined geometry with no resolved boundaries; returned geometries may well be invalid

If y is missing, *st_union(x)* returns a single geometry with resolved boundaries, else the geometries for all unioned pairs of x\[i\] and y\[j\]

*st_union()* is used to derive the coastal outline sf tibble data.frame, sg_sf.

```{r}
sg_sf <- mpsz_sf %>%
  st_union()

plot(sg_sf)

sg_sf_combine <- mpsz_sf %>%
  st_combine()

plot(sg_sf_combine)
```

## Issue 3: as.SpatialGridDataFrame.im()

The *as.SpatialGridDataFrame.im()* function can only be used if **maptools** is installed

```{r}
#| eval: false

# gridded_kde_childcareSG_ad <- maptools::as.SpatialGridDataFrame.im(kde_childcareSG_adaptive)

# spplot(gridded_kde_childcareSG_ad)
```

Alternatively, we can use the *as()* function.

```{r}
# gridded_kde_childcareSG_ad <- as(kde_childcareSG_adaptive, "SpatialGridDataFrame")
```

## Geospatial Analytics for Social Good: Myanmar Armed Conflict Case Study

Import ACLED data.

```{r}

acled_sf <- read_csv("data/ACLED_Myanmar.csv") %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(crs = 32647) %>%
  mutate(event_date = dmy(event_date))

acled_sf
```
