---
title: "In-Class Exercise 4"
author: "Immanuel Leong"
date: "September 9, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# In-Class Exercise 4

## Setup

Load the required packages.

```{r}
pacman::p_load(sf, raster, spatstat, sparr, tmap, tidyverse)
```

## Importing and Preparing Study Area

### Importing study area

```{r}

kbb_sf <- st_read(dsn = "data/rawdata", 
               layer = "Kepulauan_Bangka_Belitung") %>%
  st_union() %>%
  st_zm(drop = TRUE, what = "ZM") %>%
  st_transform(crs = 32748)

# POLYGON Z: defined in 3 dimensions (includes height)
```

### Convert into OWIN

Convert kbb_sf into an owin object.

```{r}
kbb_owin <- as.owin(kbb_sf)
kbb_owin
```

Confirm that the output is indeed an owin object.

```{r}
class(kbb_owin)
```

## Importing and Preparing Forest Fire data

Next, we import the forest fire data set.

```{r}
fire_sf <- read_csv("data/rawdata/forestfires.csv")  %>%
  st_as_sf(coords = c("longitude", "latitude"),
           crs = 4326) %>%
  st_transform(crs = 32748)

class(fire_sf)
```

Because ppp object only accepts numerical or character, the code chunk below is used to convert data from acq_date to numeric.

```{r}
fire_sf <- fire_sf %>%
  mutate(DayofYear = yday(acq_date)) %>%
  mutate(Month_num = month(acq_date)) %>%
  mutate(Month_fac = month(acq_date, label = TRUE, abbr = FALSE))
```

We prepare a map of the fire points.

```{r}
tm_shape(kbb_sf) +
  tm_polygons() +
  tm_shape(fire_sf) +
  tm_dots()
```

We prepare a point symbol map with monthly geographic distribution of forest fires in 2023.

```{r}
tm_shape(kbb_sf) +
  tm_polygons() +
  tm_shape(fire_sf) +
  tm_dots(size = 0.1) +
  tm_facets(by = "Month_fac", free.coords = FALSE, drop.units = TRUE)

# free.coords = FALSE ensures constant zoom level/map extent
```

## Computing SKDE by Month

Remove unwanted fields from fire_sf as as.ppp() only requires the mark field and geometry field from the input.

```{r}
fire_month <- fire_sf %>%
  dplyr::select(Month_num)
```

```{r}
fire_month_ppp <- as.ppp(fire_month)

fire_month_ppp
```

We will check if there are duplicated point events using the code below.

```{r}
any(duplicated(fire_month_ppp))
```

Combine the ppp and owin objects.

```{r}
fire_month_owin <- fire_month_ppp[kbb_owin]
```

### Calculating the Spatio-temporal KDE

```{r}
st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)
```

### Plotting the STKDE

```{r}
#| fig-width: 12
#| fig-height: 10

tims <- c(7,8,9,10,11,12)
  
par(mfcol = c(2,3))

for (i in tims) {
  plot(st_kde, i,
       override.par = FALSE,
       fix.range = TRUE,
       main = paste("KDE at month", i))
  
}
```

### Alternative Method

We can analyse this continuously instead.

```{r, eval = FALSE}

kde_yday <- spattemp.density(
  fire_yday_owin
)

summary(kde_yday)
```
